\newpage
{\Huge \bf Acknowledgements}

\vspace{24pt} 
First and foremost, I would like to thank my supervisors, Dr Steven Hand and Dr
Anil Madhavapeddy, for their guidance and stimulation throughout the course of
this work. Dr Hand was instrumental in transforming a rough idea into a
full-fledged framework. His emphasis on a sound evaluation motivated me to look
at every system artifact from the point of view of an eventual experiment. Dr
Madhavapeddy was helpful in setting up the experimental environment and
providing insight into operating system, virtualization and network stack
internals.

This work would be nowhere without the insight provided into CIEL internals by
Malte Schwarzkopf, Derek Murray, and Chris Smowton. In addition, Dr Kenneth
Birman (Cornell) deserves a mention for his help and patience while I wrestled
with Isis2 and Mono.

Furthermore, friends at Cambridge who have metamorphosed my sojourn here into a
tour de force  -- socially, academically, and intellectually -- merit my special
gratitude. Especially, Natalie Dobson for being a constant source of inspiration
and energy. My coursemates in the MPhil Advanced Computer Science have also
never ceased to amaze me with their aptitude and support. In the same vein,
everyone in the Computer Laboratory -- from the amazing teachers and researchers
to the extremely invaluable staff -- deserves a thank you for their
encouragement and help.

Last, but by no means least, I would like to thank my family who form the rock
on which I balance my existence. This dissertation is dedicated to them.

\vfill

\newpage
\vspace*{\fill}
